A data-driven, pixel-level model to improve the precision of \Kepler\ photometry

### 2013-10-11 documents/plm.tex
authors and affils

@@

\newcommand{\Kepler}{\textsl{Kepler}}
\newcommand{\name}{PLM}

@@

\begin{abstract}
The precision of \Kepler\ photometry%
  ---the most precise photometric measurements of stars ever made---%
  appears to be limited by unknown or untracked variations
  in spacecraft pointing, point-spread function, stellar motions and parallaxes, and temperature.
Here we present \name,
  a data-driven model of these variations in \Kepler\ data (or really of their impact on the photometry).
Importantly (and uniquely),
  \name\ works at the pixel level (not the photometric measurement level);
  it can capture more fine-grained information about the variation of the spacecraft than is available in the photometry.
\name\ is extreme in that it has enormous flexibility and it only uses data (and meta-data) to model data:
\name\ provides a prediction for each readout pixel built from a linear combination of the readouts
  from very large numbers of nearby pixels (plus some spacecraft information);
  the choice of nearby pixels is guided by ideas from causal inference.
\name\ avoids over-fitting by employing a train-and-test formalism
  designed to ensure that transit-like photometric events and short-timescale stellar variability
  cannot be captured or distorted by the model;
  it is designed to remove spacecraft-induced variability but not intrinsic stellar variability.
We show that \name\ outperforms the \Kepler\ Presearch Data Conditioning method on a set of example stars.
We release open-source code that provides \name\ output for any star in the existing \Kepler\ Archive,
  and we discuss applications for other missions,
  including any possible \Kepler\ two-wheel mission.
\end{abstract}

@@

One method that is designed to model spacecraft-induced problems
  but \emph{not} interfere with measurements of stellar variability
  is the \Kepler\ Presearch Data Conditioning (PDC) method (CITE).
The PDC ``de-trends’’ the \Kepler\ lightcurves by fitting them with a small set of basis lightcurves
  generated from a principal components analysis (PCA) of filtered lightcurves.
That is, it uses data to model data,
  regularizing the fit (and avoiding over-fitting) by filtering and restricting the dimensionality (through PCA).
The method proposed here, \name, is similar in spirit to the PDC.
The main differences are
  (1)~that it works in the pixel domain, not the lightcurve domain, so it has access to more fine-grained information,
  (2)~that it has far more freedom (far more parameters)
  but it strictly avoids over-fitting the lightcurves on exoplanet-transit time-scales through a train-and-test framework, and
  (3)~that when it fits a star, there is no possibility of any contribution of the star itself to the fit basis vectors.

>>>>

% This file is part of the causal-kepler project
% Copyright 2013 the authors.

@@

\newcommand{\notenglish}[1]{\textit{#1}}
\newcommand{\sic}{\notenglish{sic}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\Kepler}{\project{Kepler}}
\newcommand{\name}{KPLM}

@@

\title{%
  A data-driven, pixel-level model to improve the precision of \Kepler\ photometry%
}
\author{%
  Dan~Foreman-Mackey\altaffilmark{\ref{CCPP}},
  David~W.~Hogg\altaffilmark{\ref{CCPP},\ref{MPIA},\ref{email}},
  Bernhard~Sch\"olkopf\altaffilmark{\ref{MPIIS}},
  others%
}

\newcounter{address}
\setcounter{address}{1}
\altaffiltext{\theaddress}{\stepcounter{address}\label{CCPP}%
  Center for Cosmology and Particle Physics, Department of Physics, New York University}
\altaffiltext{\theaddress}{\stepcounter{address}\label{MPIA}%
  Max-Planck-Institut f\"ur Astronomie, Heidelberg, Germany}
\altaffiltext{\theaddress}{\stepcounter{address}\label{email}%
  To whom correspondence should be addressed; \texttt{<david.hogg@nyu.edu>}.}
\altaffiltext{\theaddress}{\stepcounter{address}\label{MPIIS}%
  Max-Planck-Institut f\"ur Intelligente Systeme, T\"ubingen}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Oxford}%
%%   Department of Physics, Oxford University}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Ames}%
%%   NASA Ames Research Center}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{CfA}%
%%   Harvard--Smithsonian Center for Astrophysics}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Courant}%
%%   Courant Institute of Mathematical Sciences, New York University}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{UCL}%
%%   Department of Physics and Astronomy, University College London}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{CMU}%
%%   McWilliams Center for Cosmology, Carnegie Mellon University}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Caltech}%
%%   Department of Astronomy, California Institute of Technology}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Columbia}%
%%   Department of Astronomy, Columbia University}

\begin{abstract}
The precision of \Kepler\ photometry%
  ---the most precise photometric measurements of stars ever made---%
  appears to be limited by unknown or untracked variations
  in spacecraft pointing, point-spread function, stellar motions and parallaxes, and temperature.
Here we present \name,
  a data-driven model of these variations in \Kepler\ data (or really of their impact on the photometry).
Importantly (and uniquely),
  \name\ works at the pixel level (not the photometric measurement level);
  it can capture more fine-grained information about the variation of the spacecraft
  (especially regarding pointing and point-spread function)
  than is available in the pixel-summed photometry.
\name\ is extreme in that it has enormous flexibility and it only uses data (and meta-data) to model data:
\name\ provides a prediction for each readout pixel built from a linear combination of the readouts
  from very large numbers of nearby pixels (plus some spacecraft information);
  the choice of nearby pixels is guided by ideas from causal inference.
\name\ avoids over-fitting by employing a train-and-test formalism
  designed to ensure that transit-like photometric events and short-timescale stellar variability
  cannot be captured or distorted by the model;
  it is designed to remove spacecraft-induced variability but not intrinsic stellar variability.
We show that \name\ outperforms the \Kepler\ Presearch (\sic) Data Conditioning method on a set of example stars.
We release open-source code that provides \name\ output for any star in the existing \Kepler\ Archive,
  and we discuss applications for other missions,
  including any possible \Kepler\ two-wheel mission.
\end{abstract}

@@

Reminder:  \Kepler\ performs photometry very naively...

One method that is designed to model spacecraft-induced problems
  but \emph{not} interfere with measurements of stellar variability
  is the \Kepler\ Presearch (\sic) Data Conditioning (PDC) method (CITE).
The PDC ``de-trends'' the \Kepler\ lightcurves by fitting them with a small set of basis lightcurves
  generated from a principal components analysis (PCA) of filtered lightcurves.
That is, it uses data to model data,
  regularizing the fit (and avoiding over-fitting) by filtering and restricting the dimensionality (through PCA).
The method proposed here, \name, is similar in spirit to the PDC.
The main differences are
  (1)~that it works in the pixel domain, not the lightcurve domain, so it has access to more fine-grained information,
  (2)~that it has far more freedom (far more parameters)
  but it strictly avoids over-fitting the lightcurves on exoplanet-transit time-scales
  through strong regularization and a train-and-test framework, and
  (3)~that when it fits a star, there is no possibility of any contribution of the star itself to the fit basis vectors.
The two methods are similar however,
  in that they both make the assumption that whatever spacecraft effects are imprinting variability on a stellar lightcurve
  must also be imprinting similar or related variability on other lightcurves.
By going to the pixel level, \name\ makes it easier for the model to capture variability
  that is coming through variations in the point-spread function
  or spacecraft roll or pointing.

\section{Generalities: Causality and data-driven models}

\section{Specifics: PLM and it's hyper-parameters}

\section{Examples and results}

\section{Discussion}
